### 2021-12-01 getdata.Rmd
correct data

---
title: "Get population and temperature data"
author: "Göran Broström"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(eha)
library(skum)
options(show.signif.stars = FALSE)
g <- function(x) (log(1 + 365 * x))^3
```

# Population data

We use the observation file rather than the  population file, so we can get 
correct parish information.

```{r readobs}
bb <- obs$id[obs$starttyp == 2]
births <- obs[obs$id %in% bb, 
    c("id", "sex", "birthdate", "nofrs", "sluttyp", "socStatus", "hisclass", 
      "enter", "exit", "ort", "ortnmn")]
births$event <- as.numeric(births$sluttyp == 2)
## NEW (and necessary):
extend <- with(births, (exit <= enter) & (event == 1))
births$exit[extend] <- births$enter[extend] + 1 / (4 * 365)
## NOW, remove all zero-length intervals:
births <- births[births$exit > births$enter, ]
## End NEW, now OK to select:
births <- age.window(births, c(0, 1))
births <- births[order(births$id, births$enter), ]
##indx <- tapply(births$id, births$id)
##births$mexit <- tapply(births$exit, births$id, max)[indx]
##births$mevent <- tapply(births$event, births$id, max)[indx]
##births <- births[!duplicated(births$id), ]
### Deaths:
did <- unique(births$id[births$event == 1])
deaths <- births[births$id %in% did, ]
indx <- tapply(deaths$id, deaths$id)
deaths$menter <- tapply(deaths$enter, deaths$id, min)[indx]
deaths <- deaths[!duplicated(deaths$id), ]
deaths$enter <- deaths$menter
deaths$menter <- NULL
deaths$enter <- 0 ## satisfied....
### Survivors:
survs <- births[!(births$id %in% did), ]
indx <- tapply(survs$id, survs$id)
survs$mexit <- tapply(survs$exit, survs$id, max)[indx]
survs <- survs[!duplicated(survs$id), ]
survs$exit <- survs$mexit
survs$mexit <- NULL
survs$enter <- 0 ## Ok ...
### Now, give birth to NEW births:
births <- rbind(deaths, survs)
births <- births[births$birthdate > 1895, ]

indx <- match(births$id, pers$id)
births$ab <- pers$ab[indx]
##births$df <- pers$df[indx]
births$illeg <- factor(births$ab == 2, labels = c("no", "yes"))
births$ab <- NULL
births$sluttyp <- NULL
births$parity <- pers$paritet_g[indx]
summary(births)                                
```

We put on 'subreg':

```{r subreg}
births$subreg <- factor(births$nofrs)
levels(births$subreg) <- c(rep("ume", 2), rep("bjur", 4), rep("sten", 3))
## Remove "sten":
births <- births[births$subreg != "sten", ]
births$subreg <- factor(births$subreg) # Remove level "sten"
```

# Neonatal mortality, an overview

```{r neo, fig.cap = "Neonatal mortality by year, all regions."}
neo <- age.window(births, c(0, 1 / 13))
surv <- as.matrix(neo[, c("enter", "exit", "event", "birthdate")])
pall <- perstat(surv, period = 1900:1951)$intensity
plot(1900:1950, pall, type = "b", ylim = c(0, 0.5), col = "blue", xlab = "Year",
     ylab = "Mortality")
abline(h = 0)
abline(h = 0.3, lty = 3)
```

## By region

```{r neoby, fig.cap = "Neonatal mortality by year  and region."}
usurv <- as.matrix(neo[neo$subreg == "ume", c("enter", "exit", "event", "birthdate")])
pume <-  perstat(usurv, period = 1900:1951)$intensity
plot(1900:1950, pall, type = "b", ylim = c(0, 0.7), col = "blue", xlab = "Year",
     ylab = "Mortality")
bsurv <- as.matrix(neo[neo$subreg == "bjur", c("enter", "exit", "event", "birthdate")])
pjur <-  perstat(bsurv, period = 1900:1951)$intensity
lines(1900:1950, pjur, type = "b", ylim = c(0, 0.5), col = "red")
##ssurv <- as.matrix(neo[neo$subreg == "sten", c("enter", "exit", "event", "birthdate")])
##sjur <-  perstat(ssurv, period = 1900:1951)$intensity
##lines(1900:1950, sjur, type = "b", ylim = c(0, 0.5), col = "black")
abline(h = 0)
abline(h = 0.3, lty = 3)
```

# Post-neonatal mortality, an overview

```{r pnneo, fig.cap = "Post-eonatal mortality by year, all regions."}
pneo <- age.window(births, c(1 / 13, 1))
surv <- as.matrix(pneo[, c("enter", "exit", "event", "birthdate")])
pall <- perstat(surv, period = 1900:1951)$intensity
plot(1900:1950, pall, type = "b", ylim = c(0, 0.1), col = "blue",
     xlab = "Year", ylab = "Mortality")
abline(h = 0)
abline(h = 0.3, lty = 3)
```

## By region

```{r pneoby, fig.cap = "Post-neonatal mortality by year and regions."}
usurv <- as.matrix(pneo[pneo$subreg == "ume", c("enter", "exit", "event", "birthdate")])
pume <-  perstat(usurv, period = 1900:1951)$intensity
plot(1900:1950, pall, type = "b", ylim = c(0, 0.1), col = "blue",
     xlab = "Year", ylab = "Mortality")
bsurv <- as.matrix(pneo[pneo$subreg == "bjur", c("enter", "exit", "event", "birthdate")])
pjur <-  perstat(bsurv, period = 1900:1951)$intensity
lines(1900:1950, pjur, type = "b", ylim = c(0, 0.5), col = "red")
##ssurv <- as.matrix(pneo[pneo$subreg == "sten", c("enter", "exit", "event", "birthdate")])
##sjur <-  perstat(ssurv, period = 1900:1951)$intensity
##lines(1900:1950, sjur, type = "b", ylim = c(0, 0.5), col = "black")
abline(h = 0)
abline(h = 0.3, lty = 3)
```

# Temperature data

We have two stations, *Umeå* and *Bjuröklubb*.

```{r gettemp}
source("R/tempdat.R")
umetemp <- tempdat("ume", c(1895, 1950))
bjurtemp <- tempdat("bjur", c(1895, 1950))
##stentemp <- tempdat("~/Forskning/Data/sten_temp.csv", c(1899, 1952))
saveRDS(umetemp, file = "data/umetemp.rds")
saveRDS(bjurtemp, file = "data/bjurtemp.rds")
```

# Make communal

```{r makecommunal}
##source("R/make_comm.R")
lagg <- 3 / 365  ## NOTE!!
##lagg <- 0
## Ume:
bume <- births[births$subreg == "ume", ]
bume <- make.communal(bume, umetemp["heat"], start = 1895, period = 1/52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
bume <- make.communal(bume, umetemp["cold"], start = 1895, period = 1 / 52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
bume <- make.communal(bume, umetemp["week"], start = 1895, period = 1 / 52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
### Ske:
bskel <- births[births$subreg == "bjur", ]
bskel <- make.communal(bskel, bjurtemp["heat"], start = 1895, period = 1/52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
bskel <- make.communal(bskel, bjurtemp["cold"], start = 1895, period = 1 / 52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
bskel <- make.communal(bskel, bjurtemp["week"], start = 1895, period = 1 / 52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
##bsten <- births[births$subreg == "sten", ]
##bsten <- make_comm(bsten, stentemp, start = 1899, period = 1/52, lag = lagg)
bume$region <- "ume"
bskel$region <- "ske"
infdat <- rbind(bume, bskel)
infdat <- infdat[order(infdat$id, infdat$enter), c("region", "id", "sex", "enter", "exit", "event", 
                                                   "birthdate", "socStatus", "hisclass", "illeg",
                                                   "parity", "heat", "cold", "week")]
saveRDS(infdat, file = "data/infdat.rds")
```

Done!