### 2021-12-07 Test/bysex.Rmd
first commit

---
title: "Female advantage in infant survival, Umeå and Skellefteå 1901-1950"
author: "Göran Broström"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: no
    toc_depth: 2
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
bibliography: ../bio.bib
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
abstract: "Is the female advantage in infant survival constant over time? Investigation of Umeå and Skellefteå 1901-1950. The answer is 'yes'."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
dt <- function(x, y) as.numeric(difftime(x, y, unit = "days")) # wrapper of 'difftime'
```

# Introduction

The following presentation relates to infants *born* in the parish(es) we investigate.
They are the parishes with codes 82750, 82780 (`ume`) and 82980, 82981, 82983, 82988 (`ske`).

The **R** code generating the output is given throughout, it gets a little bit messy, maybe.
I am using the **R** [@cran] package *eha* [@ehar2; @eha] for the analyses.

It is all about creating tables of number of *deaths* and total *exposure* time by *sex* and *year*.
And running *Poisson regression* on the tables.

# Get data

Data are extracted from *DDB-uttaget U15014* [@u15014], which is transformed into an **R** package *skum*, available on request.

```{r readdata}
library(skum)
births.id <- obs$id[obs$starttyp == 2] # ID of Born in parish
births <- obs[obs$id %in% births.id, # Records of born in parish
              c("id", "sex", "foddat", "startdat", "slutdat", 
                "sluttyp", "nofrs")]
births$enter <- with(births, dt(startdat, foddat)) # Tid i dagar
births$exit <- with(births, dt(slutdat, foddat))
births$event <- births$sluttyp == 2 # event == death
## Fix deaths at foddat: (Problem: age at death? 6 hours?)
fix <- (births$startdat >= births$slutdat) & births$event
births$exit[fix] <- births$enter[fix] + 6 / 24 # Add 6 hours
births <- births[births$enter < births$exit, ] # Remove zero-length intervals
infants <- age.window(births, c(0, 365)) # Cut at age 365 days
infants$birthdate <- toTime(infants$foddat)
## Remove 'intermediates':
infants <- infants[order(infants$id, -infants$exit), ]
infants <- infants[!duplicated(infants$id), ]
infants$enter <- 0 # All born in parish!
infants$sluttyp <- infants$startdat <- infants$slutdat <- NULL # Not needed any more.
infants$exit <- infants$exit / 365 # Measure age in years.
infants <- cal.window(infants, c(1901, 1951)) # Select the years 1901-1950.
infants <- infants[order(infants$birthdate), ] # Sort by birthdate
## Define coastal areas 'ume' and 'ske':
infants$region <- factor(infants$nofrs)
levels(infants$region) <- c(rep("ume", 2), rep("ske", 4), 
                            rep("inland", 3))
infants$nofrs <- NULL # Not needed any more
## Remove region 'inland':
infants <- infants[infants$region != "inland", ]
infants$region <- factor(infants$region) # Removes unused level 'inland'
source("../R/tbl.R")
x <- head(infants, 5) # First 5 lines of 'infants', store in 'x'
tbl(x, caption = "First lines of the data frame 'infants', Umeå and Skellefteå.",
    fs = 11) # fs = 'font size'
```

```{r lastlines}
y <- tail(infants, 5) # Last 5 lines of 'infants'.
tbl(y, caption = "Last lines of the data frame 'infants'.", fs = 11)
```

# Table by year and sex

See Table \@ref(tab:tabs) for the result of the following code.

```{r tabs}
## Period statistics (period = year)
girls <- with(infants[infants$sex == "female", ], 
              perstat(cbind(enter, exit, event, birthdate), 
                      period = 1901:1951, age = c(0, 1)))
boys <- with(infants[infants$sex == "male", ], 
              perstat(cbind(enter, exit, event, birthdate), 
                      period = 1901:1951, age = c(0, 1)))
## Rearrange:
boy <- data.frame(year = 1901:1950, sex = rep("boys", 50),
                  deaths = as.numeric(boys$events), 
                  exposure = as.numeric(boys$exposure))

girl <- data.frame(year = 1901:1950, sex = rep("girls", 50),
                   deaths = as.numeric(girls$events), 
                   exposure = as.numeric(girls$exposure))
## Put together
infant <- rbind(boy, girl)
infant <- infant[order(infant$year, infant$sex), ]
infant$sex <- as.factor(infant$sex)
infant$year <- as.factor(infant$year)
z <- head(infant, 5)
tbl(z, caption = "Number of deaths and total exposure time (years) by sex and year. Early years.")
```

```{r lateyrs}
w <- tail(infant, 5)
tbl(w, caption = "Number of deaths and total exposure time (years) by sex and year. Late years.")
```

# Simple analysis

Is the female survival advantage constant over 50 years 1901--1950? A simple *Poisson regression*
gives the answer:

```{r poisreg}
infant$age <- "0-1"
fit <- tpchreg(oe(deaths, exposure) ~ sex * year, time = age, data = infant)
drop1(fit, test = "Chisq") # Likelihood Ratio Test
```

Obviously, there is no evidence of that the female advantage should vary with time.
The overall femal advantage is found by the multiplicative model, see Table \@ref(tab:mult).

```{r mult, results = 'asis'}
fit <- tpchreg(oe(deaths, exposure) ~ sex + as.numeric(year),
               time = age, data = infant)
x <- summary(fit)
hr <-  fit$coefficients[1]
eupper <- round(exp(hr + 1.96 * sqrt(fit$var[1,1])), 3)
elower <- round(exp(hr - 1.96 * sqrt(fit$var[1,1])), 3)
ehr <- round(exp(hr), 3)
ltx(fit, caption = "Infant mortality by year and sex.", label = "tab:mult")
```

So, the female infant mortality is estimated to `r 100 * ehr` per cent of the male, 
with 95% confidence interval (`r format(100 * elower, nsmall = 1)`, `r 100 * eupper`) per cent.

# Graph by year

```{r byyear, fig.cap = "Female mortality in relation to male (blue squares) by year with 95 per cent pointwise confidence intervals."}
par(las = 1)
hr <- numeric(50)
up <- numeric(50)
low <- numeric(50)
i <- 0
for (year in 1901:1950){
    i <- i + 1
    me <- infant[infant$year == year, ]
    ff <- tpchreg(oe(deaths, exposure) ~ sex, time = age,
                  data = me)
    beta <- ff$coefficients[1]
    sd <- sqrt(ff$var[1, 1])
    hr[i] <- exp(beta)
    up[i] <- exp(beta + 1.96 * sd)
    low[i] <- exp(beta - 1.96 * sd)
}
plot(1901:1950, hr, type = "p", ylim = c(0.3, 2.3), 
     xlab = "Year", ylab = "Hazard ratio (per cent)", 
     col = "blue", axes = FALSE, cex = 5, pch = ".")
axis(1, at = c(1901, 1913, 1925, 1938, 1950))
axis(2, at = c(0.5, ehr, 1, 1.5, 2), 
     labels = c(0.5, ehr, 1, 1.5, 2) * 100)
box()
for (i in 1:50){
    lines(c(1900 + i, 1900 + i), c(low[i], up[i]), 
          lty = 3, col = "darkgreen")
    points(c(1900 + i, 1900 + i), c(low[i], up[i]), lty = 3, 
           col = "darkgreen", pch = 5, cex = 0.3)
}
abline(h = 1)
abline(h = ehr, lty = 2, col = "blue")
abline(h = c(eupper, elower), lty = 3, col = "blue")
```

The numbers of deaths decrease rapidly with time, see Figure \@ref(fig:derap),
which explains the increasing lenghts of the confidence intervals with time.

```{r derap, fig.cap = "Infant mortality by year, 1901-1950."}
deaths <- with(infant, tapply(deaths, year, sum))
exposures <- with(infant, tapply(exposure, year, sum))
intensity <- deaths / exposures
aver <- sum(deaths) / sum(exposures)
plot(1901:1950, intensity, type = "b", axes = FALSE, 
     ylim = c(0, 0.13),  xlab = "Year", 
     ylab = "Mortality, per cent", col = "blue")
axis(1, at = c(1901, 1913, 1925, 1938, 1950))
axis(2, at = seq(0.00, 0.12, by = 0.02), 
     labels = seq(0, 12, by = 2), las = 1)
box()
abline(h = 0)
abline(h = aver, col = "blue", lty = 2)
text(1942, 0.08, "average", col = "blue")
```

# Conclusion

Nothing much happens, the variation over the years we can see in the relative 
difference in mortality between boys and girls is obviously purely random, no
trend is visible. That is, the female advantage is strong and persistent.

One obvious observation, on the other hand, is that infant mortality
decreases markedly over the first half of the 20th century.