### 2021-12-22 getdata3.Rmd
further development based on skum::pers

---
title: "getdata3.Rmd"
author: "Göran Broström"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(skum) # Also loads 'eha'
```

# Introduction

Now (December 22, 2021) I have come to the conclusion that data from the data frame 
*persons* is the best and easiest start.

So we start with *skum::persons* instead of *skum::obs*, and age is to begin with in days,
with stillbirths age *removed* and born alive but dead on birthday at age 0.25 days.

For user of *eha::make.communal*, we need to temporarily change time unit to years.

Further, we eventually skip the use of eha::make.communal in  favor of *match*. Reason: 
eha::make.communal assumes age is measured in *years*. But this is no promise yet.

# Read skum::pers

```{r readpers}
per <- pers[(pers$frsbobtyp == 2) & (pers$df == 0), ]
per$frsbobtyp <- per$df <- per$f2id <- per$m2id <- NULL
per$dodors <- per$frsbobdatmin <- per$frsbobdatmax <- per$utdatmax <- per$utdatmin <- NULL
per <- per[(per$foddat > 18950000), ]
dn <- per$doddat < 1
per$doddat[dn] <- per$foddat[dn] + 40000
## Fix foddat:
yr <- per$foddat %% 10000 == 0
per$foddat[yr] <- per$foddat[yr] + 101 # As early as possible
mon <- per$foddat %% 100 == 0
per$foddat[mon] <- per$foddat[mon] + 1 # Ditto
## Fix doddat:
yr <- per$doddat %% 10000 == 0
per$doddat[yr] <- per$doddat[yr] + 1231 # As late as possible
mon <- per$doddat %% 100 == 0
per$doddat[mon] <- per$doddat[mon] + 28 # Ditto (guard for February ...)
##
per$birthdate <- as.Date(as.character(per$foddat), format = "%Y%m%d")
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
nabir <- is.na(per$birthdate)
per$foddat[nabir] <- per$foddat[nabir] - 2
per$doddat[nabir] <- per$doddat[nabir] - 2
##
per$birthdate <- as.Date(as.character(per$foddat), format = "%Y%m%d")
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate))
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
## Fix doddat:
per$doddat[is.na(per$deathdate)] <- per$doddat[is.na(per$deathdate)] - 1
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
##
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate, units = "days"))
oj <- per$exit < 0
per$doddat[oj] <- per$doddat[oj] + 10000
per$deathdate <-  as.Date(as.character(per$doddat), format = "%Y%m%d")
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate, units = "days"))
per$event <- per$exit <= 365
per$exit[per$exit == 0] <- 0.25
per$exit[per$exit > 365] <- 365
```

Now we need *nofrs* for newborn, found in the *observation* file(s), so

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
per <- per[(per$subreg != "sten"), ]
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))

##saveRDS(per, file = "data/infdat3.rds")
```

# Put on temperature data

```{r makecommunal}
temp_start <- 1894
temp_end <- 1951
temp_ivl <- c(temp_start, temp_end)
##
umetemp <- readRDS("data/umetemp.rds")
bjurtemp <- readRDS("data/bjurtemp.rds")
##

lagg <- 3 / 365  ## NOTE!!
tempvars <-  c("heat", "aver", "cold", "emintemp", "emeantemp", "emaxtemp", 
               "week", "year") 
## Ume:
bume <- per[per$subreg == "ume", ]
bume <- make.communal(bume, umetemp[, tempvars], 
                      start = temp_start, period = 1/52, lag = lagg)

### Ske:
bskel <- per[per$subreg == "bjur", ]
bskel <- make.communal(bskel, bjurtemp[, tempvars], 
                       start = temp_start, period = 1/52, lag = lagg)

bume$region <- "ume"
bskel$region <- "ske"
infdat <- rbind(bume, bskel)
infdat <- infdat[order(infdat$id, infdat$enter), c("region", "id", "sex", "enter", "exit", "event", 
                                                   "birthdate", "socStatus", "hisclass", "illeg",
                                                   "parity", tempvars)]
saveRDS(infdat, file = "data/infdat2_year2.rds")

```