### 2022-01-15 README.Rmd
first commit

---
title: Infant mortality, climate and weather
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Paper for presentation in Madrid, March 2022

The file *paper.pdf* is the result of this project.