### 2022-02-02 README.Rmd
updates

---
title: Infant mortality, climate and weather
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Paper for presentation in Madrid, March 2022

The file *paper3.pdf* is the result of this project. Note the time stamp on the front
page, the report is updated regularly for some time.