### 2022-02-02 ts.Rmd
first commit

---
title: "Time series"
author: "Göran Broström"
date: "`Sys.time()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Temperature

## Umeå

```{r readdataume}
umetemp <- readRDS("data/umetemp.rds")
sketemp <- readRDS("data/bjurtemp.rds")
stentemp <- readRDS("data/stentemp.rds")
par(cex = 0.8)
for (yr in 1894:1951){
    with(umetemp[umetemp$year == yr, ], plot(week, aver, 
                                               type = "b", col = "blue",
                                               main = yr,
                                               xlab = "Week", lwd = 1.5,
                                               ylim = c(-15, 15), axes = FALSE))
    axis(1, at = c(1, seq(4,52, by = 4)))
    axis(2)
    box()
    ##with(umetemp[umetemp$year == yr, ], lines(week, emeantemp, type = "l"))
    abline (h = 0)
    abline(h = c(-5, -10, -15), v = c(1, seq(4, 52, by = 4)), lty = 3)
    with(sketemp[sketemp$year == yr, ], lines(week, aver, type = "b", col = "red"))
    with(stentemp[stentemp$year == yr, ], lines(week, aver, type = "b", col = "black"))
    legend("topright", legend = c("Umeå", "Bjuröklubb", "Stensele"), col = c("blue", "red", "black"),
           lty = 1, pch = "o", cex = 1)
}
```

# Deaths

## Umeå

```{r deathsume}
infdat <- readRDS("data/infdat3_year3.rds")
umeinf <- infdat[infdat$subreg == "ume", ]
```

## Skellefteå

```{r deathsske}
skeinf <- infdat[infdat$subreg == "bjur", ]
```