### 2022-02-07 README.Rmd
updated, BP weq

---
title: Infant mortality, climate and weather
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Paper for presentation in Madrid, March 2022

Now Bourgeois-Pichat and the log-cube transform are erased, but the present state of
*paper4.pdf* chaotic. No text is updated, only some text is removed.

The file *paper3.pdf* is the old result of the project. Note the time stamp on the front
page, this report is *not* updated for some time.